biclaR: Estimating the socio-environmental impacts of car substitution by bicycle and public transit using open tools

### 2023-11-27 paper/PaperCEUS/PaperCEUS.Rmd
replace tables paper

For the cycling legs of the journey (first and last legs), socio-environmental impacts were estimated, using the HEAT for Cycling tool v5.0 [@HEAT] from the World Health Organization, and the [`HEATaaS` R package](https://github.com/HEAT-WHO/HEAT_heatr_api)[^5].
The HEAT tool provided estimates on the shifting from car to cycling for a short term time horizon (i.e., one year) and the long term (i.e., ten years).
We considered two dimensions: *social* — including the physical activity of cyclists, air pollution exposure, and road casualties; *environmental* — including CO~2~eq emissions and other pollutants.

@@

Table \ref{tab:summary1} presents the LMA total daily trips that can be made with cycling + TP combination (with the aforementioned restrictions), the trips in the baseline scenario and corresponding new trips to achieve the national strategy targets (4% and 10%), for different route profiles. 
For the cycling legs of the journey (first and last legs), the environmental avoided emissions and monetized socio-environment (SE) benefits are also presented, resulting from replacing car trips with cycling.

```{r summary1}
summary1 = readRDS("load/summary1.Rds")

summary1 = summary1 %>%
  mutate(LTS = recode(LTS, `3` = "safe", `4` = "direct")) |> 
  mutate(Estrategia = paste0(Estrategia*100,"%")) |> 
  mutate(TotalTrips = TotalTrips/2,
         Bikebase = Bikebase/2,
         Bikepotential = Bikepotential/2) #this is a quick correction while we donr run biclar entirely again. The error was found and it was adding up the 2 bike legs, instead of making the mean. The SE estimates remain the same because they are from the bikeshare and carshare percentages, that did not change.

knitr::kable(summary1,
      digits = 0,
      format.args = list(big.mark = " "),
      col.names = c("Target", "Routing", "Total trips",
                    "Baseline Cycling + PT", "Potencial Cycling + PT",
                    "Avoided CO2eq (ton/yr)", "SE Benefits for 10 years (thousand €)"),
      caption = "\\label{summary1}Summary of the cycling potencial of intermodality scenario and its socio-environmental benefits for the cycling legs.",
      format="latex",
      booktabs = TRUE) |> 
  kableExtra::column_spec(column = c(4:7), width = "6em")

@@

```{r summary22}
summary22 = readRDS("load/TPtable_aml.Rds")
summary22 = summary22 %>%
  mutate(LTS = recode(Percursos, "seguro" = "safe", "direto" = "direct")) |> 
  mutate(Value = money_emissions_eur/1000) |> 
  select(Meta_ENMAC, LTS, CO2eq_TP, CO_PT, PM10_PT, NOx_PT, VOC_PT, Value) |> 
  arrange(Meta_ENMAC)

@@

```{r summaryall, message=FALSE, warning=FALSE}
summary0 = readRDS("load/HEAT_aml_results.Rds") |> slice_tail(n = 4)
rate = ((1+0.03)^10)/((1+0.05)^10) # inflation rate 3% discount rate 5%

summaryall = summary1 |> 
  mutate(Meta_ENMAC = Estrategia) |>
  left_join(summary22) |> 
  mutate(Code = c(3304, 3404, 3310, 3410)) |> 
  left_join(summary0 |> select(Code, Economic), by = "Code") |> 
  select(Estrategia, LTS, co2, CO2eq_TP, monet, Value, Economic) |> 
  mutate(CO2eq = co2 + CO2eq_TP,
         value1 = Economic/1000 + Value, # this is CORRECT
         value10 = monet + Value*10*rate) |> # this is CORRECT
  select(Estrategia, LTS, CO2eq, value1, value10)

>>>>

For the cycling legs of the journey (first and last legs), socio-environmental impacts were estimated, using the HEAT for Cycling tool v5.2 [@HEAT] from the World Health Organization, and the [`HEATaaS` R package](https://github.com/HEAT-WHO/HEAT_heatr_api)[^5].
The HEAT tool provided estimates on the shifting from car to cycling for a short term time horizon (i.e., one year) and the long term (i.e., ten years).
We considered two dimensions: *social* — including the physical activity of cyclists, air pollution exposure, and road casualties; *environmental* — including CO~2~eq emissions and other pollutants.

@@

Table \ref{tab:summary1} presents the LMA total daily trips that can be made with cycling + TP combination (with the aforementioned restrictions), the trips in the baseline scenario and corresponding new trips to achieve the national strategy targets (4% and 10%), for different route profiles. 
For the cycling legs of the journey (first and last legs), the environmental avoided emissions and monetized socio-environment (SE) benefits are presented in \ref{tab:summary1b}, resulting from replacing car trips with cycling.

```{r summary0}
summary0 = readRDS(url("https://github.com/U-Shift/biclar/releases/download/0.0.1/HEAT_AML_intermodal_ppp_NEW.Rds"))
# see code/test-code/heat-prep.R#intermodality

summary0 = summary0 |> 
    mutate(Target = paste0(ENMAC*100,"%"),
           Routing = c("safe", "safe", "direct", "direct"),
           TotalTrips = Total,
           Bikebase = Bike,
           Bikepotential = Bike_new,
           Social = Economic - Environmenal_only,
           Social10 = Economic10 - Environmenal10_only) 
```

```{r summary1}
summary1 = summary0 |> 
  select(Target, Routing, TotalTrips, Bikebase, Bikepotential) |> 
    arrange(desc(Target))

knitr::kable(summary1,
      digits = 0,
      format.args = list(big.mark = " "),
      col.names = c("Target", "Routing", "Total trips",
                    "Baseline Cycling + PT", "Potencial Cycling + PT"),
      caption = "\\label{summary1}Summary of the cycling potencial of the intermodality scenario.",
      format="latex",
      booktabs = TRUE) |> 
  kableExtra::column_spec(column = c(4:5), width = "7em")

```{r summary1b}
summary1b = summary0 |> 
  select(Target, Routing, Mortality, Social, CO2eq, Environmenal_only) |> 
  mutate(Social = Social/1000,
         Environmenal_only = Environmenal_only/1000) |> 
    arrange(desc(Target))

knitr::kable(summary1b,
      digits = c(0,0,1,0,0,0),
      format.args = list(big.mark = " "),
      col.names = c("Target", "Routing",
                    "Avoided Mortality (deaths/yr)", "Social benefits (k€/yr)", 
                    "Avoided CO2eq (ton/yr)", "Environmental benefits (k€/yr)"),
      caption = "\\label{summary1b}Summary of the cycling potencial of intermodality scenario and its socio-environmental benefits for the cycling legs.",
      format="latex",
      booktabs = TRUE) |> 
  kableExtra::column_spec(column = c(3:6), width = "6em")

@@

```{r summary22}
summary22 = readRDS("load/TPtable_aml.Rds")
summary22 = summary22 %>%
  mutate(Routing = recode(Percursos, "seguro" = "safe", "direto" = "direct")) |> 
  mutate(Value = money_emissions_eur/1000,
         Target = Meta_ENMAC) |> 
  select(Target, Routing, CO2eq_TP, CO_PT, PM10_PT, NOx_PT, VOC_PT, Value) |> 
  arrange(Target)

@@

```{r summaryall, message=FALSE, warning=FALSE}

rate = ((1+0.03)^10)/((1+0.05)^10) # inflation rate 3% discount rate 5%

summaryall = summary0 |> 
  left_join(summary22) |> 
  select(Target, Routing, CO2eq, CO2eq_TP, Environmenal_only, Environmenal10_only, Social, Social10, Value) |> 
    mutate(Social = Social/1000,
           Social10 = Social10/1000,
         Environmenal_only = Environmenal_only/1000,
         Environmenal10_only = Environmenal10_only/1000) |> 
  mutate(CO2eqT = CO2eq + CO2eq_TP,
         value1 = Environmenal_only + Value + Social, # this is CORRECT
         value10 = Environmenal10_only + Value*10*rate + Social10) |> # this is CORRECT
  select(Target, Routing, CO2eqT, value1, value10)

### 2023-11-27 paper/PaperCEUS/PaperCEUS.tex
replace tables paper

For the cycling legs of the journey (first and last legs),
socio-environmental impacts were estimated, using the HEAT for Cycling
tool v5.0 (Kahlmeier et al., 2017) from the World Health Organization,
and the
\href{https://github.com/HEAT-WHO/HEAT_heatr_api}{\texttt{HEATaaS} R
package}\footnote{\texttt{HEATaaS} is under development. For more
  information contact
  \href{https://heatwalkingcycling.org}{heatwalkingcycling.org}.}. The
HEAT tool provided estimates on the shifting from car to cycling for a
short term time horizon (i.e., one year) and the long term (i.e., ten
years). We considered two dimensions: \emph{social} --- including the
physical activity of cyclists, air pollution exposure, and road
casualties; \emph{environmental} --- including CO\textsubscript{2}eq
emissions and other pollutants.

@@

Table \ref{tab:summary1} presents the LMA total daily trips that can be
made with cycling + TP combination (with the aforementioned
restrictions), the trips in the baseline scenario and corresponding new
trips to achieve the national strategy targets (4\% and 10\%), for
different route profiles. For the cycling legs of the journey (first and
last legs), the environmental avoided emissions and monetized
socio-environment (SE) benefits are also presented, resulting from
replacing car trips with cycling.

@@

\caption{\label{tab:summary1}\label{summary1}Summary of the cycling potencial of intermodality scenario and its socio-environmental benefits for the cycling legs.}
\centering
\begin{tabular}[t]{llr>{\raggedleft\arraybackslash}p{6em}>{\raggedleft\arraybackslash}p{6em}>{\raggedleft\arraybackslash}p{6em}>{\raggedleft\arraybackslash}p{6em}}
\toprule
Target & Routing & Total trips & Baseline Cycling + PT & Potencial Cycling + PT & Avoided CO2eq (ton/yr) & SE Benefits for 10 years (thousand €)\\
\midrule
4\% & safe & 538 514 & 2 312 & 20 385 & 5 920 & 230 270\\
4\% & direct & 500 880 & 2 274 & 18 944 & 6 011 & 223 720\\
10\% & safe & 538 514 & 2 312 & 52 324 & 15 192 & 591 790\\
10\% & direct & 500 880 & 2 274 & 48 609 & 15 414 & 574 200\\
\bottomrule
\end{tabular}
\end{table}

@@

\caption{\label{tab:summaryall}\label{summaryall}Summary of the avoided CO2eq emmissions (ton/year) and the estimated social and environmental benefits (monetized in thousand €) by replacing car trips with cycling in combination with PT.}
\centering
\begin{tabular}[t]{ll>{\raggedleft\arraybackslash}p{8em}>{\raggedleft\arraybackslash}p{8em}>{\raggedleft\arraybackslash}p{8em}}
\toprule
Target & Routing & Avoided CO2eq (tons) & SE Benefits 1yr (k€) & SE Benefits 10yrs (k€)\\
\midrule
4\% & safe & 14 513 & 27 299 & 242 024\\
4\% & direct & 14 713 & 26 572 & 235 706\\
10\% & safe & 35 819 & 69 938 & 620 094\\
10\% & direct & 36 207 & 67 956 & 602 972\\
\bottomrule
\end{tabular}
\end{table}

>>>>

For the cycling legs of the journey (first and last legs),
socio-environmental impacts were estimated, using the HEAT for Cycling
tool v5.2 (Kahlmeier et al., 2017) from the World Health Organization,
and the
\href{https://github.com/HEAT-WHO/HEAT_heatr_api}{\texttt{HEATaaS} R
package}\footnote{\texttt{HEATaaS} is under development. For more
  information contact
  \href{https://heatwalkingcycling.org}{heatwalkingcycling.org}.}. The
HEAT tool provided estimates on the shifting from car to cycling for a
short term time horizon (i.e., one year) and the long term (i.e., ten
years). We considered two dimensions: \emph{social} --- including the
physical activity of cyclists, air pollution exposure, and road
casualties; \emph{environmental} --- including CO\textsubscript{2}eq
emissions and other pollutants.

@@

Table \ref{tab:summary1} presents the LMA total daily trips that can be
made with cycling + TP combination (with the aforementioned
restrictions), the trips in the baseline scenario and corresponding new
trips to achieve the national strategy targets (4\% and 10\%), for
different route profiles. For the cycling legs of the journey (first and
last legs), the environmental avoided emissions and monetized
socio-environment (SE) benefits are presented in \ref{tab:summary1b},
resulting from replacing car trips with cycling.

@@

\caption{\label{tab:summary1}\label{summary1}Summary of the cycling potencial of the intermodality scenario.}
\centering
\begin{tabular}[t]{llr>{\raggedleft\arraybackslash}p{7em}>{\raggedleft\arraybackslash}p{7em}}
\toprule
Target & Routing & Total trips & Baseline Cycling + PT & Potencial Cycling + PT\\
\midrule
4\% & safe & 538 514 & 2 312 & 20 385\\
4\% & direct & 500 880 & 2 274 & 18 944\\
10\% & safe & 538 514 & 2 312 & 52 323\\
10\% & direct & 500 880 & 2 274 & 48 609\\
\bottomrule
\end{tabular}
\end{table}

\caption{\label{tab:summary1b}\label{summary1b}Summary of the cycling potencial of intermodality scenario and its socio-environmental benefits for the cycling legs.}
\centering
\begin{tabular}[t]{ll>{\raggedleft\arraybackslash}p{6em}>{\raggedleft\arraybackslash}p{6em}>{\raggedleft\arraybackslash}p{6em}>{\raggedleft\arraybackslash}p{6em}}
\toprule
Target & Routing & Avoided Mortality (deaths/yr) & Social benefits (k€/yr) & Avoided CO2eq (ton/yr) & Environmental benefits (k€/yr)\\
\midrule
4\% & safe & 4.1 & 12 717 & 2 958 & 238\\
4\% & direct & 4.0 & 12 441 & 3 004 & 241\\
10\% & safe & 10.0 & 32 820 & 7 590 & 610\\
10\% & direct & 10.0 & 31 800 & 7 694 & 618\\
\bottomrule
\end{tabular}
\end{table}

@@

\caption{\label{tab:summaryall}\label{summaryall}Summary of the avoided CO2eq emmissions (ton/year) and the estimated social and environmental benefits (monetized in thousand €) by replacing car trips with cycling in combination with PT.}
\centering
\begin{tabular}[t]{ll>{\raggedleft\arraybackslash}p{8em}>{\raggedleft\arraybackslash}p{8em}>{\raggedleft\arraybackslash}p{8em}}
\toprule
Target & Routing & Avoided CO2eq (tons) & SE Benefits 1yr (k€) & SE Benefits 10yrs (k€)\\
\midrule
4\% & safe & 11 551 & 14 380 & 127 534\\
10\% & safe & 28 217 & 36 861 & 325 814\\
4\% & direct & 11 706 & 14 135 & 125 016\\
10\% & direct & 28 487 & 35 905 & 318 062\\
\bottomrule
\end{tabular}
\end{table}
